---
title: "Week 6 Exercise"
author: "Jacques Uber"
date: "5/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is my solution to the problem 9.44:
_For each of the following problems, use the eight-step approach to test the hypotheses_:

$$a) \quad H_{0}: p = .28 \quad H_{a}: p > .28 \quad n = 783 \quad x = 230 \quad \alpha = 0.1$$
$$b) \quad H_{0}: p = .61 \quad H_{a}: p \ne .61 \quad n = 401  \quad \hat{p} = .56 \quad \alpha = 0.05$$

# The Eight Steps of Testing Hypotheses
The eight step approach is:

1. Establish a null and alternative hypothesis.

2. Determine the appropriate statistical test.

3. Set the value of alpha, the Type 1 error rate.

4. Establish the decision rule.

5. Gather sample data.

6. Analyze the data

7. Reach a statistical conclusion

8. Make a business decision

## Part a
Steps 1, 3, and 5 are done for us. For step 2, we see that the alternative is testing p greater than .28, so this means we are looking for a one tail test. We can use a z score to see if we can solve for $\hat{p}$ which we _can_ do by our known values of $x$, $n$, and the formula:

$$\hat{p} = \frac{x}{n}$$

We will reject our null hypothesis if (step 4) our z score is greater than $z_{\alpha}$.

```{r}
alpha = 0.1
x = 230
n = 783
p = .28
q = 1 - p
phat = x/n
z = (phat - p)/sqrt((p*q)/n)
"Step 7. We should reject the null hypothesis?"
z > pnorm(alpha, 0, 1)
```

Step 8 is impossible because we are note provided context for the numbers in the problem.

## Part b
Steps 1, 3, and 5 are done for us. For step 2, we see that we are looking for non-equality which means we need to do a two tail test. Again, we can use a zscore but this time we will reject the null when our z value to less than $-z_{\alpha/2}$ or greater than $z_{\alpha/2}$.

```{r}
phat = .56
alpha = 0.05
n = 401
p = 0.61
q = 1 - p
z = (phat - p)/sqrt((p*q)/n)
"Step 7. We should reject the null hypothesis"
z < -pnorm(alpha/2, 0, 1) || z > pnorm(alpha/2, 0, 1)
```

Again, step 8 is impossible because we are not provided context.