---
title: "Week 2 Excercise"
author: "Jacques Uber"
date: "4/9/2020"
output:
  pdf_document: default
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```
Here is the solution to 3.40.

## Define the data
First we load 2010 U.S. Census data for a random sample of 40 households where each value represents the age of the first person recorded for that household.

```{r}
data <- c(
  42, 29, 31, 38, 55, 27, 28,
  33, 49, 70, 25, 21, 38, 47,
  63, 22, 38, 52, 50, 41, 19,
  22, 29, 81, 52, 26, 35, 38,
  29, 31, 48, 26, 33, 42, 58,
  40, 32, 24, 34, 25
)
```

The problem asks us to calculate $P_{30}$, $P_{80}$, $Q_{1}$, $Q_{3}$, the range, and the interquartile range for our data.

## Define a helper function
R provides a `quantile(..., type=2)` function that can easily help us calculate the values we need, but for the sake of my own learning, lets define our own function using the definition provided in Black on page 50. 

```{r}

myQuartile = function (xs, P) {
  # xs is assumed to be a vector
  # P is the percentile we are trying to fine
  xs = sort(xs)
  i = (P / 100) * length(xs)
  if (i %% 1 == 0) {
    # i is a whole number
    if (i + 1 >= length(xs)) {
      result = xs[i]
    } else {
      result = (xs[i] + xs[i + 1]) / 2
    }
  } else {
    # i is not a whole number
    result = xs[floor(i + 1)]
  }
  
  return(result)
}
```

## Solution
$P_{30}$ = `r myQuartile(data, 30)` (check answer `r quantile(data, 0.30, type=2)`)

$P_{80}$ = `r myQuartile(data, 80)` (check answer `r quantile(data, 0.80, type=2)`)

$Q_{1}$ = `r myQuartile(data, 25)` (check answer `r quantile(data, 0.25, type=2)`)

$Q_{3}$ = `r myQuartile(data, 75)` (check answer `r quantile(data, 0.75, type=2)`)

$range$ = `r max(data) - min(data)`

$IQR$ = $Q{3} - Q{1}$ = `r myQuartile(data, 75) - myQuartile(data, 25)`

The code used to calculate this was:
```
$P_{30}$ = myQuartile(data, 30) (check answer quantile(data, 0.30, type=2))

$P_{80}$ = myQuartile(data, 80) (check answer quantile(data, 0.80, type=2))

$Q_{1}$ = myQuartile(data, 25) (check answer quantile(data, 0.25, type=2))

$Q_{3}$ = myQuartile(data, 75) (check answer quantile(data, 0.75, type=2))

$range$ = max(data) - min(data)

$IQR$ = $Q{3} - Q{1}$ = r myQuartile(data, 75) - myQuartile(data, 25)
```