---
title: "Week 5 Practice"
author: "Jacques Uber"
date: "5/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library(readxl)
```

# Problem
Chapter 8 Analyzing the Databases # 1

_Using the manufacturing database as a sample, construct a 95% confidence interval for the population mean number of production workers. What is the point estimate? How much is the error of the estimate? Comment on the results._
```{r}
BLACK_DATABASES = "./data/Databases_in_Excel.xlsx"
data = read_excel(BLACK_DATABASES, sheet = "Manufacturer")
```

We don't know the population mean or the population standard deviation, so its likely that a t-distribution is our best option. What does our sample population look like?
```{r, warning=FALSE}
library(ggplot2)
ggplot(data, aes(x=`No. Prod. Wkrs.`)) + geom_histogram()
```
Unfortunately, this doesn't look normal -- it is skewed heavily to the right. According to Black, though, the t-distribution is a relatively robust statistic so the fact that our population distribution is not normal should be okay.

Lets now compute our 95% confidence interval using the formula:

$$\overline{x} - t_{\alpha/2,n-1} \frac{s}{\sqrt{n}} \leq \mu \leq \overline{x} + t_{\alpha/2,n-1} \frac{s}{\sqrt{n}} $$

```{r}
sample_data = data %>% sample_frac(1)
n = nrow(sample_data)
df = n - 1
alpha = 0.95
xbar = mean(sample_data$`No. Prod. Wkrs.`)
s = sd(sample_data$`No. Prod. Wkrs.`)
t = dt(alpha/2, df)
error = t * (s/sqrt(n))
low = xbar - error
high = xbar + error
```


```{r}
sprintf("%s <= population average <= %s with %s precent confidence", round(low,2), round(high, 2), round(100*alpha,0))
sprintf("The point estimate is %s (which is also the population mean", round(xbar, 2))
sprintf("The error is %s", round(error, 2))
```

I'm not confident I approached this problem correctly.