\documentclass[12pt]{article}
\usepackage[hmargin=1in, vmargin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{setspace}
\pagestyle{fancy}
\usepackage[small]{caption}
\usepackage{lastpage}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{url}
\DeclareGraphicsExtensions{.png}

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}
\definecolor{grey}{gray}{0.9}


\lstset{
    tabsize=2,
    breaklines=true,
    showstringspaces=false
    backgroundcolor=\color{grey},
    basicstyle=\ttfamily\scriptsize,
}


\def\author{Jacques Uber}
\def\title{Contributing to OSS: Final Report}
\def\date{\today}

\fancyhf{} % clear all header and footer fields
\fancyhead[LO]{\author}
\fancyhead[RO]{\date}
\renewcommand{\headrulewidth}{0pt}
% The weird spacing here is to get the spacing of \thepage to be right.
\fancyfoot[C]{\thepage\
                    / 1}

\setcounter{secnumdepth}{0}
\setlength{\parindent}{0pt}
\setlength{\parskip}{4mm}
\linespread{1}
% Talk about how the hinge works and where the power switch it relative to how the hinge swings.
% Pictures:
%   Use copyright: Name, Year
% Get rid of value language
%
% OO: Wednesday 9-12
%
\begin{document}
%\fancyhead[CO]{\title}
\begin{center}
\underline{
\large{\title}
}
\end{center}
\singlespacing

\section{Final Report}
\subsection{What worked?}
Just using the software on an almost daily basis gave me a lot of exposure to
the code. Subscribing to the mailing list was also helpful because it gave me a
view of the project that I would normally not see as just a typical user.
Filtering through tickets looking for issues to work on was also helpful, as I
now feel like I have a better view of what is being worked on.
\subsection{What did not work?}
I did not find the strategy of purposely seeking out issues to work. It may be
because the project I chose (Django) is a very well maintained and active
project; there is not a lot of low hanging fruit. The best contribution I made
was when I was using the project's code and found a bug.
\subsection{What you have learned?}
I learned that opening a ticket is not as scary as it sounds.
\subsection{What is the top advice you wish someone had given you at the start of this course?}
For the purpose of contributing for the first time, choose a project that
doesn't have so many contributers and has a lot of low hanging fruit. You may
need to initially shop around for projects, but it will make the assignments
go more smoothly.
\subsection{How will you continue your involvement with FOSS in the future, if at all?}
Because I have now done it, I will find it easier to open bug reports and
fix issues while using Django.
\end{document}
